---
title: "Figure 1 Metzincin protease paper"
author: "Jasmin Wächter"
date: "2/9/2022"
output: html_document
---

```{r}

setwd("/Volumes/FREECOM HDD/R_ADAM")


library(ape)
library(DoubletFinder)
library(remotes)
library(Matrix)
library(fields)
library(KernSmooth)
library(ROCR)
library(parallel)
library(SingleCellExperiment)
library(Seurat)
library(tidyverse)
library(Matrix)
library(scales)
library(cowplot)
library(RCurl)
library(dplyr)
library(stringr)
library(ggplot2)
library(fields)
library(clustree)
library(ggraph)
library(plotly)
library(MAST)
library(multtest)
library(metap)

library(SeuratWrappers)
library(VGAM)
library(patchwork)
library(RColorBrewer)
library(plotly)



#Load the trophoblast object generated by Matthew Shannon
load("/Volumes/FREECOM HDD/R_ADAM/Trophoblast.Rdata")

```

#Add clustern names independently
```{r}
Trophoblasts@meta.data$cluster_name<- NA

Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^0"))] <- "CTB 1"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^1"))] <- "CTB 2"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^2"))] <- "CTB 3"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^3"))] <- "CTB 4"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$CellType, "^4"))] <- "cCTB 1"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^7"))] <- "cCTB 2"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^5"))] <- "EVT"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^6"))] <- "SCTp"

```

#Reorder clusters
```{r}
Idents(Trophoblasts) <- "cluster_name"
#this is to have the clusters appear in the order that you desire
my_levels <- c("SCTp", "CTB 1","CTB 2", "CTB 3", "CTB 4", "cCTB 1","cCTB 2", "EVT")

# Re-level object@ident
Trophoblasts@active.ident <- factor(x = Trophoblasts@active.ident, levels = my_levels)
#View(Trophoblasts@meta.data)

```

#Figure 1A
```{r}
pdf("metzincin_cluster_heatmap.pdf",  width=6, height=5)
Idents(Trophoblasts) <- "cluster_name"
UMAPPlot     (Trophoblasts, 
              pt.size = 1,
              
              label= TRUE,
              shuffle = FALSE)
dev.off()


jpeg("metzincin_cluster_heatmap2.jpeg",  width=6, height=5, units="in", res=300)
UMAPPlot     (Trophoblasts, 
              pt.size = 1,
              label= FALSE)+NoLegend()
dev.off()
```

#Individual cluster colours for fig1A
```{r}
jpeg("EVT_cluster_heatmap.pdf",  width=6, height=5, res=300, units="in")

UMAPPlot     (Trophoblasts, 
             cols = c('0' = 'lightgrey',
                                     "1"= "lightgrey",
                                     "2" = "lightgrey",
                                     '3' = 'lightgrey',
                                     '4'    = 'lightgrey',
                                     '5'   = '#FF61CC',
                                     '6'   = 'lightgrey',
                                     "7" = 'lightgrey'),

              pt.size = 1,
              label= FALSE)+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())

dev.off()

jpeg("CTB_cluster_heatmap.pdf",  width=6, height=5, res=300, units="in")
UMAPPlot     (Trophoblasts, 
             cols = c('0' = '#CD9600',
                                     "1"= "#7CAE00",
                                     "2" = "#00BE67",
                                     '3' = '#00BFC4',
                                     '4'    = 'lightgrey',
                                     '5'   = 'lightgrey',
                                     '6'   = 'lightgrey',
                                     "7" = 'lightgrey'),

              pt.size = 1,
              label= FALSE)+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())

dev.off()

jpeg("SCT_cluster_heatmap.pdf",  width=6, height=5, res=300, units="in")

UMAPPlot     (Trophoblasts, 
             cols = c('0' = 'lightgrey',
                                     "1"= "lightgrey",
                                     "2" = "lightgrey",
                                     '3' = 'lightgrey',
                                     '4'    = 'lightgrey',
                                     '5'   = 'lightgrey',
                                     '6'   = '#F8766D',
                                     "7" = 'lightgrey'),

              pt.size = 1,
              label= FALSE)+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())

dev.off()

jpeg("cCTB_cluster_heatmap.pdf",  width=6, height=5, res=300, units="in")
UMAPPlot     (Trophoblasts, 
             cols = c('0' = 'lightgrey',
                                     "1"= "lightgrey",
                                     "2" = "lightgrey",
                                     '3' = 'lightgrey',
                                     '4'    = '#00A9FF',
                                     '5'   = 'lightgrey',
                                     '6'   = 'lightgrey',
                                     "7" = '#C77CFF'),

              pt.size = 1,
              label= FALSE)+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()






```

#VLn plot of all relevant metzincins
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"

#MMP genes
VlnPlot  (object = Trophoblasts, features = c("MMP2"))
VlnPlot  (object = Trophoblasts, features = c("MMP3"))
VlnPlot  (object = Trophoblasts, features = c("MMP7"))
VlnPlot  (object = Trophoblasts, features = c("MMP8"))
VlnPlot  (object = Trophoblasts, features = c("MMP9"))
#VlnPlot  (object = Trophoblasts, features = c("MMP10")) - no expression
VlnPlot  (object = Trophoblasts, features = c("MMP11"))
#VlnPlot  (object = Trophoblasts, features = c("MMP12")) - not expressed 
#VlnPlot  (object = Trophoblasts, features = c("MMP13")) - no expression
VlnPlot  (object = Trophoblasts, features = c("MMP14"))
VlnPlot  (object = Trophoblasts, features = c("MMP15"))
VlnPlot  (object = Trophoblasts, features = c("MMP16"))
VlnPlot  (object = Trophoblasts, features = c("MMP17"))
VlnPlot  (object = Trophoblasts, features = c("MMP19"))
VlnPlot  (object = Trophoblasts, features = c("MMP20"))
VlnPlot  (object = Trophoblasts, features = c("MMP21"))
#VlnPlot  (object = Trophoblasts, features = c("MMP23A")) - not expressed 
VlnPlot  (object = Trophoblasts, features = c("MMP23B"))
VlnPlot  (object = Trophoblasts, features = c("MMP24"))
VlnPlot  (object = Trophoblasts, features = c("MMP25"))
VlnPlot  (object = Trophoblasts, features = c("MMP26"))
#VlnPlot  (object = Trophoblasts, features = c("MMP27")) - no expression
VlnPlot  (object = Trophoblasts, features = c("MMP28"))

#ADAMTS genes
VlnPlot  (object = Trophoblasts, features = c("ADAMTS1"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS2"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS3"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS4"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS5"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS6"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS7"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS8"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS9"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS10"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS12"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS13"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS14"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS15"))
#VlnPlot  (object = Trophoblasts, features = c("ADAMTS16")) no expression at all
VlnPlot  (object = Trophoblasts, features = c("ADAMTS17"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS18"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS19"))
VlnPlot  (object = Trophoblasts, features = c("ADAMTS20"))

#Pappalysin genes
VlnPlot  (object = Trophoblasts, features = c("PAPPA2"))
VlnPlot  (object = Trophoblasts, features = c("PAPPA"))

#Timp and RECK ADAMTSL
VlnPlot  (object = Trophoblasts, features = c("TIMP1"))
VlnPlot  (object = Trophoblasts, features = c("TIMP2"))
VlnPlot  (object = Trophoblasts, features = c("TIMP3"))
VlnPlot  (object = Trophoblasts, features = c("TIMP4"))
VlnPlot  (object = Trophoblasts, features = c("RECK"))

#ADAM genes  ( only with protease domains)
VlnPlot  (object = Trophoblasts, features = c("ADAM8"))
VlnPlot  (object = Trophoblasts, features = c("ADAM9"))
VlnPlot  (object = Trophoblasts, features = c("ADAM10"))
VlnPlot  (object = Trophoblasts, features = c("ADAM12"))
VlnPlot  (object = Trophoblasts, features = c("ADAM15"))
VlnPlot  (object = Trophoblasts, features = c("ADAM17"))
VlnPlot  (object = Trophoblasts, features = c("ADAM19"))
VlnPlot  (object = Trophoblasts, features = c("ADAM20"))
VlnPlot  (object = Trophoblasts, features = c("ADAM21"))
VlnPlot  (object = Trophoblasts, features = c("ADAM28"))
VlnPlot  (object = Trophoblasts, features = c("ADAM33"))

#ADAM genes with no catalytic domain
#VlnPlot  (object = Trophoblasts, features = c("ADAM1"))
VlnPlot  (object = Trophoblasts, features = c("ADAM2"))
#VlnPlot  (object = Trophoblasts, features = c("ADAM7")) - no expression
VlnPlot  (object = Trophoblasts, features = c("ADAM11"))
VlnPlot  (object = Trophoblasts, features = c("ADAM18"))
VlnPlot  (object = Trophoblasts, features = c("ADAM20"))
VlnPlot  (object = Trophoblasts, features = c("ADAM21"))
VlnPlot  (object = Trophoblasts, features = c("ADAM22"))
#VlnPlot  (object = Trophoblasts, features = c("ADAM23")) - no expression
VlnPlot  (object = Trophoblasts, features = c("ADAM29"))
#VlnPlot  (object = Trophoblasts, features = c("ADAM30")) - no expression 

#astacins 
VlnPlot  (object = Trophoblasts, features = c("TLL1"))
#VlnPlot  (object = Trophoblasts, features = c("ASTL")) - no expression
VlnPlot  (object = Trophoblasts, features = c("BMP1"))
VlnPlot  (object = Trophoblasts, features = c("MEP1A"))
VlnPlot  (object = Trophoblasts, features = c("MEP1B"))
VlnPlot  (object = Trophoblasts, features = c("TLL2"))


#Metzincins with zero expression were excluded from follow-up analyses. indicated by # at the beginning of the line

#The analysis will be continued with the following

All_metzincins<-c("MMP2","MMP3","MMP7","MMP8","MMP9","MMP11","MMP14","MMP15","MMP16","MMP17","MMP19","MMP20","MMP21","MMP23B","MMP24","MMP25","MMP26","MMP28","ADAMTS1","ADAMTS2","ADAMTS3","ADAMTS4","ADAMTS5","ADAMTS6","ADAMTS7","ADAMTS8","ADAMTS9","ADAMTS10","ADAMTS12","ADAMTS13","ADAMTS14","ADAMTS15","ADAMTS17","ADAMTS18","ADAMTS19","ADAMTS20","PAPPA2","PAPPA","TIMP1","TIMP2","TIMP3","TIMP4","RECK","ADAMTSL2","ADAMTSL3","ADAMTSL4","ADAMTSL5","ADAM8","ADAM9","ADAM10","ADAM12","ADAM15","ADAM17","ADAM19","ADAM20","ADAM21","ADAM28","ADAM33","ADAM2","ADAM11","ADAM18","ADAM20","ADAM21","ADAM22","ADAM29","TLL1","BMP1","MEP1A","MEP1B","TLL2")

```

#calculate the percentage of cells that are expressing certain genes
```{r}
PrctCellExpringGene <- function(object, genes, group.by = "all"){
    if(group.by == "all"){
        prct = unlist(lapply(genes,calc_helper, object=object))
        result = data.frame(Markers = genes, Cell_proportion = prct)
        return(result)
    }

    else{        
        list = SplitObject(object, group.by)
        factors = names(list)

        results = lapply(list, PrctCellExpringGene, genes=genes)
        for(i in 1:length(factors)){
        results[[i]]$Feature = factors[i]
        }
        combined = do.call("rbind", results)
        return(combined)
    }
}

calc_helper <- function(object,genes){
    counts = object[['RNA']]@counts
    ncells = ncol(counts)
    if(genes %in% row.names(counts)){
    sum(counts[genes,]>0)/ncells
    }else{return(NA)}
}


#Calculate % of gene expression
P<-PrctCellExpringGene(Trophoblasts, All_metzincins)

#Calculate # of cells the gens are expressed in - 7798 is total amount of cells in our dataset

P$cell_number<-(P$Cell_proportion*7798)

P
```

#Which markers are expressed in more than 1% of cells
```{r}

Keepers<-P[which(P$Cell_proportion>0.01),]

OnIce<-P[which(P$Cell_proportion<0.01),]

OnIce_names<-OnIce$Markers

Keepers_names<-Keepers$Markers
```

#Other thresholds such as expression in >50 cells or >100 were also investigated by heatmap and violin plot
```{r}
Invest<-P$cell_number<100
Invest<-P$cell_number<50
```



#Check expression of markers that are in between 1%. 
```{r}

Idents(Trophoblasts)<-"ID"


VlnPlot  (object = Trophoblasts, features = "MMP9")
VlnPlot  (object = Trophoblasts, features = "MMP23B")
VlnPlot  (object = Trophoblasts, features = "ADAMTS5")

VlnPlot  (object = Trophoblasts, features = "ADAMTS20")
VlnPlot  (object = Trophoblasts, features = "PAPPA")
VlnPlot  (object = Trophoblasts, features = "TIMP4")





```

#How many cells does each sample contribute 
```{r}
table(Trophoblasts$ID)

```



#Investigate in between metzincins via heatmap
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"

#MMP genes


FeaturePlot  (object = Trophoblasts, features = c("MMP9"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("MMP23B"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("PAPPA"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("ADAMTS5"),  sort.cell = TRUE, pt.size = 1,  min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("ADAMTS20"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("TIMP4"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("ADAMTSL5"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("ADAMTSL4"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("TLL1"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))

FeaturePlot  (object = Trophoblasts, features = c("TLL2"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))



```

#Investigate in between metzincins via heatmap part 2
```{r}


Idents(Trophoblasts)<-"cluster_name"


VlnPlot  (object = Trophoblasts, features = "MMP9")
VlnPlot  (object = Trophoblasts, features = "MMP23B")
VlnPlot  (object = Trophoblasts, features = "ADAMTS5")

VlnPlot  (object = Trophoblasts, features = "ADAMTS20")
VlnPlot  (object = Trophoblasts, features = "PAPPA")
VlnPlot  (object = Trophoblasts, features = "TIMP4")

VlnPlot  (object = Trophoblasts, features = "ADAMTSL4")
VlnPlot  (object = Trophoblasts, features = "ADAMTSL5")



```





#are any markers enriched in CTB
```{r}

CTB_only<- subset(Trophoblasts, idents = c("CTB 1","CTB 2","CTB 3","CTB 4"))



CTB_saviour<-PrctCellExpringGene(CTB_only_org, OnIce_names)


CTB_keepers<-CTB_saviour[which(CTB_saviour$Cell_proportion>0.05),]

CTB_Keepers_names<-CTB_keepers$Markers

rm(CTB_only_org)
```

#are any markers enriched in cCTB
```{r}

cCTB_only<- subset(Trophoblasts, idents = c("cCTB"))



cCTB_saviour<-PrctCellExpringGene(cCTB_only, OnIce_names)

cCTB_keepers<-cCTB_saviour[which(cCTB_saviour$Cell_proportion>0.05),]

cCTB_Keepers_names<-cCTB_keepers$Markers


rm(cCTB_only)
```
#are any markers enriched in SCT
```{r}

SCT_only<- subset(Trophoblasts, idents = c("SCTp"))



SCT_saviour<-PrctCellExpringGene(SCT_only, OnIce_names)

SCT_keepers<-SCT_saviour[which(SCT_saviour$Cell_proportion>0.05),]

SCT_Keepers_names<-SCT_keepers$Markers

rm(SCT_only)
```

#are any markers enriched in EVT
```{r}

EVT_only<- subset(Trophoblasts, idents = c("EVT"))



EVT_saviour<-PrctCellExpringGene(EVT_only, OnIce_names)

EVT_keepers<-EVT_saviour[which(EVT_saviour$Cell_proportion>0.05),]

EVT_Keepers_names<-EVT_keepers$Markers

rm(EVT_only)
```

#Final metzincins left
```{r}
Metzincin<-c(Keepers_names,CTB_Keepers_names,cCTB_Keepers_names,EVT_Keepers_names,SCT_Keepers_names)

#Trophoblast markers
housekeep<-c("HLA-G","ERVFRD-1","TEAD4", "CDX2","NOTCH1")
```


#Heatmap of all remaining genes - Figure 1B 
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
cluster.averages <- AverageExpression (Trophoblasts, return.seurat = TRUE)

pdf("ADAM_all_heatmap.pdf",  width=6, height=5)

DoHeatmap(cluster.averages, features = Metzincin ,
                  size         = 5,
                      angle        = 270,
                      hjust        = 1,
                      group.bar    = FALSE,
                      raster       = FALSE,
                      draw.lines   = FALSE)+
scale_fill_gradientn (colours      = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100)) +
theme                (axis.text.x  = element_text     (vjust = 1, hjust = 0.5,  face = "bold",angle = 0, colour = "black"))+
theme                (axis.text.y  = element_text     ( face = "bold", colour = "black"))+
theme                (legend.text  = element_text     (size  = 15))+
  guides( colour = FALSE)




dev.off()



```
#HOUSEKEEPING ONLY HEATMAP - Figure 1B
```{r}

DoHeatmap(cluster.averages, features = housekeep ,
                  size         = 10,
                      angle        = 270,
                      hjust        = 1,
                      group.bar    = FALSE,
                      raster       = FALSE,
                      draw.lines   = FALSE)+
scale_fill_gradientn (colours      = colorRampPalette (rev (brewer.pal (n = 7, name = "RdYlBu"))) (100)) +
theme                (axis.text.x  = element_text     (vjust = 1, hjust = 0.5,  face = "bold",angle = 0, colour = "black"))+
theme                (axis.text.y  = element_text     ( face = "bold", colour = "black"))+
theme                (legend.text  = element_text     (size  = 10))+
  guides( colour = FALSE)



```
#Figure 1C 

#CTB metzincins 
```{r}
Idents(Trophoblasts) <- "cluster_name"


pdf("Violin_ADAMTS6.pdf",  width=6, height=5)
VlnPlot  (object = Trophoblasts, features = c("ADAMTS6"), pt.size = 0)+NoLegend()
dev.off()

pdf("Violin_ADAMTS19.pdf",  width=6, height=5)
VlnPlot  (object = Trophoblasts, features = c("ADAMTS19"), pt.size = 0)+NoLegend()
dev.off()

jpeg("Feature_ADAMTS6.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("ADAMTS6"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()

jpeg("Feature_ADAMTS19.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("ADAMTS19"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()


```

#EVT metzincins 
```{r}
pdf("Violin_ADAM28.pdf",  width=6, height=5)
VlnPlot  (object = Trophoblasts, features = c("ADAM28"), pt.size = 0)+NoLegend()
dev.off()

pdf("Violin_MMP11.pdf",  width=6, height=5)
VlnPlot  (object = Trophoblasts, features = c("MMP11"), pt.size = 0)+NoLegend()
dev.off()

jpeg("Feature_ADAM28.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("ADAM28"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()

jpeg("Feature_MMP11.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("MMP11"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()

```


#Ubiquitous and terminally differentiated
```{r}


jpeg("Feature_ADAM17.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("ADAM17"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()

pdf("Violin_ADAM17.pdf",  width=6, height=5)
VlnPlot  (object = Trophoblasts, features = c("ADAM17"), pt.size = 0)+NoLegend()+ theme(axis.line = element_line(size = 1))
dev.off()

pdf("Violin_PAPPA2.pdf",  width=6, height=5)
VlnPlot  (object = Trophoblasts, features = c("PAPPA2"), pt.size = 0)+NoLegend()+ theme(axis.line = element_line(size = 1))
dev.off()

jpeg("Feature_PAPPA2.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("PAPPA2"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()

```

