---
title: "Figure 6 Metzincin protease paper"
author: "Jasmin Wächter"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
library(Seurat)
#library(SeuratWrappers)
library(nichenetr)
library                   (tidyverse)
library                   (Matrix)
library                   (dplyr)
library                   (ggplot2)
library                   (cowplot)
library                   (tibble)
library                   (readxl)
library(EnhancedVolcano)
```

#Load trophoblast object
```{r}
load("/Volumes/FREECOM HDD/R_ADAM/Trophoblast.Rdata")
```

#load the differential gene expression analysis performed for figure 5 
```{r}
load("/Volumes/FREECOM HDD/R_ADAM/data/5_DEG/CTB_markers_all25aug21.RData")
```
#Figure 6A
```{r}
EnhancedVolcano(CTB_markers,
                lab = CTB_markers$gene,
                x = "avg_log2FC",
                y = "p_val_adj",
               xlim = c(-6, 4),
 ylim =c(0,320),
 selectLab = c("FBN2"),
               col=c("grey","grey","grey","grey"))
```

#Figure 6B

```{r}
CTB_only<- subset(Trophoblasts, idents = c("CTB 1","CTB 2","CTB 3","CTB 4"))
```

```{r}
PrctCellExpringGene <- function(object, genes, group.by = "all"){
    if(group.by == "all"){
        prct = unlist(lapply(genes,calc_helper, object=object))
        result = data.frame(Markers = genes, Cell_proportion = prct)
        return(result)
    }

    else{        
        list = SplitObject(object, group.by)
        factors = names(list)

        results = lapply(list, PrctCellExpringGene, genes=genes)
        for(i in 1:length(factors)){
        results[[i]]$Feature = factors[i]
        }
        combined = do.call("rbind", results)
        return(combined)
    }
}

calc_helper <- function(object,genes){
    counts = object[['RNA']]@counts
    ncells = ncol(counts)
    if(genes %in% row.names(counts)){
    sum(counts[genes,]>0)/ncells
    }else{return(NA)}
}

```

#How many cells express ADAMTS6, FBN2 and combined
```{r}
ADAMTS6_CTB <- subset (CTB_only, subset =   ADAMTS6     > 0 )

ADAMTS6_CTB

PrctCellExpringGene(ADAMTS6_CTB, c("ADAMTS6","FBN2"))

FBN2_CTB <- subset (CTB_only, subset =   FBN2    > 0 )

FBN2_CTB


PrctCellExpringGene(FBN2_CTB, c("ADAMTS6","FBN2"))

	
```

#Figure 6C
#ADAMTS6-FBN2
```{r}
pdf("Violin_ADAMTS6_org.pdf",  width=6, height=5)
VlnPlot  (object = interface.integrated, features = c("ADAMTS6"), pt.size = 0)+NoLegend()
dev.off()

pdf("Violin_ADAMTS6_org.pdf",  width=6, height=5)
VlnPlot  (object = interface.integrated, features = c("ADAMTS6"), pt.size = 0)+NoLegend()
dev.off()

jpeg("Feature_FBN2.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = interface.integrated, features = c("FBN2"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()

jpeg("Feature_FBN2.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = interface.integrated, features = c("FBN2"),  sort.cell = TRUE, pt.size = 1,   min.cutoff = 0, order = TRUE, cols = c("grey", "red"))+NoLegend()+ theme(axis.line = element_blank(),axis.title=element_blank(),axis.text = element_blank(),axis.ticks = element_blank())
dev.off()

```


#Figure 6D

#Load previously constructed ligand target matrix and network weights
```{r}
load("/Volumes/FREECOM HDD/R_ADAM/data/4_ligand analysis/1_final objects/30_sept_ligand_target_matrix_full.Rdata")

load("/Volumes/FREECOM HDD/R_ADAM/data/4_ligand analysis/1_final objects/30sept_weighted_networks_full.Rdata")
```


```{r}
library(nichenetr)
library(tidyverse)


ligands_all = c("ADAMTS6") # this can be a list of multiple ligands if required
targets_all = c("FBN2")

View(ligand_tf_matrix)

active_signaling_network = get_ligand_signaling_path(ligand_tf_matrix = ligand_target_matrix2, ligands_all = ligands_all, targets_all = targets_all, weighted_networks = weighted_networks)

# For better visualization of edge weigths: normalize edge weights to make them comparable between signaling and gene regulatory interactions
active_signaling_network_min_max = active_signaling_network
active_signaling_network_min_max$sig = active_signaling_network_min_max$sig %>% mutate(weight = ((weight-min(weight))/(max(weight)-min(weight))) + 0.75)
active_signaling_network_min_max$gr = active_signaling_network_min_max$gr %>% mutate(weight = ((weight-min(weight))/(max(weight)-min(weight))) + 0.75)


graph_min_max = diagrammer_format_signaling_graph(signaling_graph_list = active_signaling_network_min_max, ligands_all = ligands_all, targets_all = targets_all, sig_color = "indianred", gr_color = "steelblue")

# To render the graph: uncomment following line of code

DiagrammeR::render_graph(graph_min_max, layout = "tree")

#Figure 6B
DiagrammeR::export_graph(graph_min_max, file_name = "ADAMTS6andFBN2connect.pdf", file_type = "pdf")


```



#Figure 6D - genes identified - what percentage in the CTB


```{r}
CTB_only<- subset(Trophoblasts, idents = c("CTB 1","CTB 2","CTB 3","CTB 4"))

PrctCellExpringGene(CTB_only, c("ADAMTS6","FBN2","IRAK1","REST","PELI2","RNF2","STAT3","FN1","SKIL","SMAD3"))
```
# Result 
```{r}
Description:df [10 × 2]
Markers
<chr>
Cell_proportion
<dbl>
ADAMTS6	0.4491509951			
FBN2	0.9046923498			
IRAK1	0.1130180756			
REST	0.3814131824			
PELI2	0.0009129085			
RNF2	0.2088734709			
STAT3	0.2882965127			
FN1	0.2148986672			
SKIL	0.0995070294			
SMAD3	0.0058426146	
```

#Conversion to colour code 
```{r}
#RGBPercent(0%,0%,0%) -> 100%
             
#RGBPercent(100%,100%,100%)-> 0%
             
1-0.4491509951	#	ADAMTS6	
1-0.9046923498	#FBN2		
1-0.1130180756	#IRAK1		
1-0.3814131824	#REST		
1-0.0009129085	#PELI2		
1-0.2088734709	#RNF2		
1-0.2882965127	#STAT3		
1-0.2148986672	#FN1		
1-0.0995070294	#SKIL		
1-0.0058426146	#SMAD3
```
```{r}
0.550849  #	ADAMTS6	-> 55% Hex	FF8C8C
0.09530765#FBN2		-> 10% Hex	FF1A1A
0.8869819#IRAK1		-> 89% Hex	FFE3E3
0.6185868#REST	->62%	Hex	FF9E9E
0.9990871#PELI2	->100%	Hex	FFFFFF
0.7911265#RNF2		-> 79%  Hex	FFC9C9
0.7117035	#STAT3->	71%	Hex	FFB5B5
0.7851013	#FN1		->79% Hex	FFC9C9
0.900493	#SKIL		->90% Hex	FFE6E6
0.9941574#SMAD3   ->99% Hex	FFFCFC


```

#6E
```{r, include = FALSE}
library(clusterProfiler)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(ReactomePA)
require(pathview)
```

#ReactomePA - calculate %cells expressed for each gene
```{r}
hs <- org.Hs.eg.db
my.symbols <- CTB_markers$gene
CTB_names<-AnnotationDbi::select(hs, 
       keys = my.symbols,
       columns = c( "SYMBOL","ENTREZID"),
       keytype = "SYMBOL")
library(tidyverse)

CTB_names<-CTB_names %>% rename(gene = SYMBOL)

CTB_names_ENTREZ<-CTB_names$ENTREZID

CTB_markers_del<-CTB_markers[which( CTB_markers$gene %in% CTB_names$gene),]

CTB_markers_del <- merge(CTB_markers_del,CTB_names,by="gene")

#--------------
CTB_percentage<-PrctCellExpringGene(CTB_only, CTB_markers_del$gene)

CTB_percentage<-CTB_percentage %>% rename(gene = Markers)

CTB_markers_del <- merge(CTB_markers_del,CTB_percentage,by="gene")

CTB_markers_del<-CTB_markers_del[which(CTB_markers_del$Cell_proportion!=0),]

#look at percentage expression in cells 
CTB_list_prct =CTB_markers_del[,9]

names(CTB_list_prct) = CTB_markers_del$ENTREZID

CTB_list_prct = sort (CTB_list_prct, decreasing = TRUE)
CTB_list_prct<-CTB_list_prct[which(!duplicated(names(CTB_list_prct)))]
```

#investigate pathways ADAMTS6 and FBN2 are involved in
```{r}
my.symbols <- c("FBN2","ADAMTS6")
help<-AnnotationDbi::select(hs, 
       keys = my.symbols,
       columns = c( "SYMBOL","UNIPROT","ENTREZID"),
       keytype = "SYMBOL")


Entrez_metzincin<-help$ENTREZID
x <- enrichPathway(gene=Entrez_metzincin, pvalueCutoff = 0.05, readable=TRUE)
head(x)
```

```{r}
viewPathway("Molecules associated with elastic fibres", 
            readable = TRUE, 
            foldChange = CTB_list_prct)+
        scale_color_continuous(low="grey", high="red", name = "Percentage", na.value = "#E5C494")

viewPathway("Elastic fibre formation", 
            readable = TRUE, 
            foldChange = CTB_list_prct)+
        scale_color_continuous(low="grey", high="red", name = "Percentage", na.value = "#E5C494")

ELASTIC_fibre_PRCT_general<-viewPathway("Molecules associated with elastic fibres", 
            readable = TRUE, 
            foldChange = CTB_list_prct)+
        scale_color_continuous(low="grey", high="red", name = "Percentage", na.value = "#E5C494")

ELASTIC_fibre_PRCT<-viewPathway("Elastic fibre formation", 
            readable = TRUE, 
            foldChange = CTB_list_prct)+
        scale_color_continuous(low="grey", high="red", name = "Percentage", na.value = "#E5C494")







#Fig6E Supplemental fig 4
ggsave("ECM_deg_PRCT_general.pdf",device = "pdf", plot= ELASTIC_fibre_PRCT_general , path = "/Users/jasmin/Desktop", width = 10, height = 5)

ggsave("ELASTIC_fibre_Prct.pdf",device = "pdf", plot= ELASTIC_fibre_PRCT , path = "/Users/jasmin/Desktop", width = 10, height = 5)
```


