---
title: "Figure 5 Metzincin protease paper"
author: "Jasmin Wächter"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
library(Seurat)
#library(SeuratWrappers)
library(nichenetr)
library                   (tidyverse)
library                   (Matrix)
library                   (dplyr)
library                   (ggplot2)
library                   (cowplot)
library                   (tibble)
library                   (readxl)
library(EnhancedVolcano)
```


#Load trophoblast object
```{r}
load("/Volumes/FREECOM HDD/R_ADAM/Trophoblast.Rdata")
```

#Load nichenet constructed networks
```{r}
gr_network = readRDS(url("https://zenodo.org/record/3260758/files/gr_network.rds"))
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
sig_network = readRDS(url("https://zenodo.org/record/3260758/files/signaling_network.rds"))
```

#load literature curated subtrates and interactions
```{r}
CTB_info<-read_excel("/Users/jasmin/Desktop/CTB_proteases_updated.xlsx")

as.data.frame(CTB_info)
```
#Inspect validity of list 
```{r}
CTB_info = CTB_info %>% 
    filter(from %in% geneinfo_human$symbol & to %in% geneinfo_human$symbol) # keep only interactions between genes with oficial gene symbols: optional step
```

#Add source to dataset
```{r}
# give your data source a name
CTB_info = CTB_info %>% mutate(source = "lit_review", database = "pubmed_etc") 
```

#add data to lr network
```{r}
new_lr_network = lr_network %>% bind_rows(CTB_info)
```

#Add new weights 
```{r}
new_network_weights_df = tibble(source = "lit_review", weight = 1)
new_source_weights_df = optimized_source_weights_df %>% bind_rows(new_network_weights_df)
```

#Make new network weights and ligand target matrix 
```{r}
# aggregate the individual data sources in a weighted manner to obtain a weighted integrated signaling network
weighted_networks = construct_weighted_networks(lr_network = new_lr_network, sig_network = sig_network, gr_network = gr_network, source_weights_df = new_source_weights_df)


# downweigh the importance of signaling and gene regulatory hubs - use the optimized parameters of this 
weighted_networks = apply_hub_corrections(weighted_networks = weighted_networks, lr_sig_hub = hyperparameter_list$lr_sig_hub, gr_hub = hyperparameter_list$gr_hub)


#Get a full list of ligands
ligands<-weighted_networks$lr_sig$from
ligands<-unique(ligands)
ligands<-as.list(ligands)


# Infer ligand-target regulatory potential scores based on the weighted integrated networks
ligand_target_matrix2 = construct_ligand_target_matrix(weighted_networks = weighted_networks, ligands = ligands, algorithm = "PPR", damping_factor = hyperparameter_list$damping_factor, ltf_cutoff = hyperparameter_list$ltf_cutoff)

#SAVE
save(ligand_target_matrix2, file = "/Volumes/FREECOM HDD/R_ADAM/30_sept_ligand_target_matrix_full.Rdata")
save(weighted_networks, file = "/Volumes/FREECOM HDD/R_ADAM/30sept_weighted_networks_full.Rdata")
```
#Define receiver cells and get their genes
```{r}
## receiver --> consider a gene expressed if only expressed in 20%
receiver = c("CTB 1","CTB 3","CTB 4", "CTB 2")
expressed_genes_receiver = get_expressed_genes(receiver, Trophoblasts, pct = 0.2, assay_oi = 'RNA')

background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix2)]
```

#Define sender cells and get their genes 
```{r}
#always use RNA slot - expressed in 20%
## sender
sender_celltypes =  c("CTB 1","CTB 3","CTB 4", "CTB 2")

list_expressed_genes_sender = sender_celltypes %>% unique() %>% lapply(get_expressed_genes, Trophoblasts, 0.2 , assay_oi = 'RNA') # lapply to get the expressed genes of every sender cell type separately here
expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique()
```

#Bind lr and sig network
```{r}
new_lr_sig_network = new_lr_network %>% bind_rows(sig_network)
```

#list data sources available and add your own
```{r}
addition<-data.frame("lit_review","pubmed_etc","literature","directional_ppi","protease_substrate")
names(addition)<-c("source","database","type_db","type_interaction","network")

annotation_data_sources_updated <- rbind(annotation_data_sources, addition)

annotation_data_sources_updated$source %>% unique()
```

#choose which datasource to keep
```{r}
data_sources_to_keep = annotation_data_sources_updated %>% filter( source %in% c("kegg_cytokines",
"kegg_cams",
"kegg_neuroactive",
"kegg_ecm",
"pharmacology",
"ramilowski_known",
"ppi_lr",
"ppi_l_bidir",
"ppi_bidir_r",
"ppi_bidir_bidir",
"ppi_lr_go",
"ppi_l_bidir_go",
"ppi_bidir_r_go",
"ppi_bidir_bidir_go",
"cpdb_interaction",
"cpdb_complex",
"omnipath_directional",
"omnipath_undirectional",
"inweb_inbio_interaction",
"inweb_inbio_interaction_pathway",
"inweb_inbio_pathway",
"pathwaycommons_interacts_with",
"pathwaycommons_in_complex_with",
"vinayagam_ppi","lit_review")) %>% pull(source)


new_lr_sig_network_strict = new_lr_sig_network %>% filter(source %in% data_sources_to_keep) 

new_lr_sig_network_strict_met<-new_lr_sig_network_strict%>%filter(from %in% c("ADAM17","ADAM10","ADAMTS1","ADAMTS6","MMP14","MMP15"))

new_lr_sig_network_personal = new_lr_sig_network %>% filter(source %in% "lit_review") 


```

#Select expressed ligands and receptors 
```{r}
ligands = new_lr_sig_network_strict %>% pull(from) %>% unique()

receptors = rownames(ligand_target_matrix2) 

expressed_ligands = intersect(ligands,expressed_genes_sender)
expressed_receptors = intersect(receptors,expressed_genes_receiver)

CTB_proteases<-c("ADAM17","ADAM10","ADAMTS1","ADAMTS6","MMP14","MMP15")

```

#Ligand - receptor interactions in trophoblast - for chosen proteases
```{r}
# get the ligand-receptor network of the top-ranked ligands
lr_network_top = new_lr_sig_network_strict_met %>% filter(from %in% CTB_proteases & to %in% expressed_receptors) %>% distinct(from,to)
best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()

lr_network_top_df = weighted_networks$lr_sig %>% filter(from %in% CTB_proteases & to %in% best_upstream_receptors)

# convert to a matrix
lr_network_top_df = lr_network_top_df %>% spread("from","weight",fill = 0)
lr_network_top_matrix = lr_network_top_df %>% select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df$to)

# perform hierarchical clustering to order the ligands and receptors
dist_receptors = dist(lr_network_top_matrix, method = "binary")
hclust_receptors = hclust(dist_receptors, method = "ward.D2")
order_receptors = hclust_receptors$labels[hclust_receptors$order]

dist_ligands = dist(lr_network_top_matrix %>% t(), method = "binary")
hclust_ligands = hclust(dist_ligands, method = "ward.D2")
order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]

vis_ligand_receptor_network = lr_network_top_matrix[order_receptors, order_ligands_receptor]
p_ligand_receptor_network = vis_ligand_receptor_network %>% t() %>% make_heatmap_ggplot("Expressed metzincin proteases","Metzincin substrates and interactions", color = "mediumvioletred", x_axis_position = "top",legend_title = "Prior interaction potential")
p_ligand_receptor_network

#Fig 5A
ggsave("CTB_all_ligand_with_combined_weights_strict",device = "pdf", plot= p_ligand_receptor_network, path = "/Users/jasmin/Desktop", width = 10, height = 5)
```

#Figure 5B gene ontology  
```{r, include = FALSE}
library(clusterProfiler)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(ggplot2)



```

#Show which pathways identified genes are involved in
```{r}
ego4 <- enrichGO(gene         = twenty,
                OrgDb         = org.Hs.eg.db,
                keyType       = 'SYMBOL',
                ont           = "CC",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.05)


goplot(ego4)


 
pdf("GOpathway_diff_expressed_genes_goplot.pdf", height = 5, width = 9)  
print(goplot(ego4))
dev.off()

```

#Figure 5C - separate substrate or interaction

```{r}
twenty<-order_receptors
choices<-new_lr_sig_network_strict_met[which( new_lr_sig_network_strict_met$to %in% twenty),] 
```

#Load information describing whether it is a substrate or an interaction
```{r}
SuborInt<-read_excel("/Users/jasmin/Desktop/Substrate_or_interaction.xlsx")

as.data.frame(SuborInt)

Interaction<-SuborInt$Interaction
Substrate<-SuborInt$Substrate
```

#Get nichenet curated interactions
```{r}
nicheInt<-choices[which(!choices$source %in% "lit_review"),]
Interaction_nichenet<-nicheInt$to %>% unique()
```

#make a CTB only meta.data slot
```{r}
Trophoblasts$CTB<- NA
 
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^CTB 1"))] <- "CTB"
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^CTB 2"))] <- "CTB"
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^CTB 3"))] <- "CTB"
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^CTB 4"))] <- "CTB"
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^cCTB 1"))] <- "non_CTB"
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^EVT"))] <- "non_CTB"
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^SCTp"))] <- "non_CTB"
Trophoblasts$CTB[which(stringr::str_detect(Trophoblasts$cluster_name, "^cCTB 2"))] <- "non_CTB"

Idents(Trophoblasts) <- "CTB"

```

#find markers expressed in the CTB cluster - DGE
```{r}
CTB_markers<- FindMarkers(Trophoblasts, 
                          ident.1 = "CTB",
                          logfc.threshold = -Inf, 
                          min.pct = -Inf, 
                          min.diff.pct = -Inf, 
                          test.use = "MAST")


CTB_markers <- CTB_markers %>% 
                rownames_to_column(var="gene") 

```
#Fig 5C
```{r}
Interactions<-CTB_markers[which( CTB_markers$gene %in% Interaction),]
Substrates<-CTB_markers[which( CTB_markers$gene %in% Substrate),]
Nichenet_interactions<-CTB_markers[which( CTB_markers$gene %in% Interaction_nichenet),]


pdf("DEG_interactions.pdf",  width=7, height=5)
EnhancedVolcano(Interactions,
                lab = Interactions$gene,
                x = "avg_log2FC",
                y = "p_val_adj",
               xlim = c(-6, 4),
               ylim =c(0,320),
               col=c("green","green","green","green")
               
       ) 
dev.off()

pdf("DEG_substrates.pdf",  width=7, height=5)
EnhancedVolcano(Substrates,
                lab = Substrates$gene,
                x = "avg_log2FC",
                y = "p_val_adj",
               xlim = c(-6, 4),
ylim =c(0,320),
               col=c("blue","blue","blue","blue")
               
       ) 

dev.off()

pdf("DEG_nichenet_interactions.pdf",  width=7, height=5)
EnhancedVolcano(Nichenet_interactions,
                lab = Nichenet_interactions$gene,
                x = "avg_log2FC",
                y = "p_val_adj",
               xlim = c(-6, 4),
               ylim =c(0,320),
               col=c("pink","pink","pink","pink")
               
       ) 

dev.off()

pdf("DEG_general.pdf",  width=7, height=5)

EnhancedVolcano(CTB_markers,
                lab = CTB_markers$gene,
                x = "avg_log2FC",
                y = "p_val_adj",
               xlim = c(-6, 4),
 ylim =c(0,320),
 selectLab = c("GGCT"),
               col=c("grey","grey","grey","grey"))

dev.off()

```
#5D
```{r}
IntandsubandNiche<-c(Interaction, Substrate, Interaction_nichenet)

IntandsubandNiche

Top_identified<-CTB_markers[which(CTB_markers$avg_log2FC>0.5 & CTB_markers$gene %in% IntandsubandNiche & CTB_markers$p_val< 10e-4),]

Top_identified

Top_proteases<-choices[which(choices$to %in% Top_identified$gene),]

Top_proteases


```




