---
title: "Supplemental figures Metzincin protease paper"
author: "Jasmin Wächter"
date: "2/9/2022"
output: html_document
---


```{r}
setwd("/Volumes/FREECOM HDD/R_ADAM")


library(ape)
library(DoubletFinder)
library(remotes)
library(Matrix)
library(fields)
library(KernSmooth)
library(ROCR)
library(parallel)
library(SingleCellExperiment)
library(Seurat)
library(tidyverse)
library(Matrix)
library(scales)
library(cowplot)
library(RCurl)
library(dplyr)
library(stringr)
library(ggplot2)
library(fields)
library(clustree)
library(ggraph)
library(plotly)
library(MAST)
library(multtest)
library(metap)

#library(monocle3)
library(monocle)
library(SeuratWrappers)
library(SCENT)
library(VGAM)
library(patchwork)
library(RColorBrewer)
library(plotly)



#From figure 1
load("/Volumes/FREECOM HDD/R_ADAM/Trophoblast.Rdata")

```

#Add clustern names independently
```{r}
Trophoblasts@meta.data$cluster_name<- NA

Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^0"))] <- "CTB 1"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^1"))] <- "CTB 2"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^2"))] <- "CTB 3"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^3"))] <- "CTB 4"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$CellType, "^4"))] <- "cCTB 1"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^7"))] <- "cCTB 2"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^5"))] <- "EVT"
Trophoblasts@meta.data$cluster_name[which(stringr::str_detect(Trophoblasts@meta.data$seurat_clusters, "^6"))] <- "SCTp"

```

#reassign cluster levels- cluster order 
```{r}
#this is to have the clusters appear in the order that you desire

# cluster re-assignment occurs, which re-assigns clustering in my_levels (assuming you have 12 clusters in total)
my_levels <- c("SCTp", "CTB 1","CTB 2", "CTB 3", "CTB 4", "cCTB 1","cCTB 2", "EVT")

# Re-level object@ident
Trophoblasts@active.ident <- factor(x = Trophoblasts@active.ident, levels = my_levels)
#View(Trophoblasts@meta.data)

```

#Supplemental Fig1 A
```{r}
pdf("ADAM_cluster_heatmap.pdf",  width=6, height=5)
Idents(Trophoblasts) <- "cluster_name"
UMAPPlot     (Trophoblasts, 
              pt.size = 1,
              
              label= TRUE,
              shuffle = FALSE)
dev.off()
View(Trophoblasts@meta.data)

jpeg("ADAM_cluster_heatmap2.jpeg",  width=6, height=5, units="in", res=300)
UMAPPlot     (Trophoblasts, 
              pt.size = 1,
              label= FALSE)+NoLegend()
dev.off()

```

#Sup Fig 1B

#TEAD4 Supplemental Figure 1
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
jpeg("TEAD4_interface.jpeg", units="in", width=6, height=5, res=300)

FeaturePlot  (object = Trophoblasts, features = c("TEAD4"),     min.cutoff = 0, order = TRUE, cols = c("grey", "red") , pt.size = 1                )
dev.off()



```

#HLA-G Supplemental Figure 1
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
jpeg("HLA_g_interface.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("MMP12"),     min.cutoff = 0, order = TRUE, cols = c("grey", "red")  , pt.size = 1               )+ NoLegend()
dev.off()
```



#ERVFRD-1 Supplemental Figure 1
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
jpeg("ERVFRD_1_interface.jpeg", units="in", width=6, height=5, res=300)
FeaturePlot  (object = Trophoblasts, features = c("ERVFRD-1"),     min.cutoff = 0, order = TRUE, cols = c("grey", "red")  , pt.size = 1               )+ NoLegend()
dev.off()
```


#EGFR Supplemental Figure 1
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
jpeg("EGFR_interface.jpeg", units="in", width=6, height=5, res=300)

FeaturePlot  (object = Trophoblasts, features = c("EGFR"),     min.cutoff = 0, order = TRUE, cols = c("grey", "red") , pt.size = 1                ) + NoLegend()
dev.off()



```

#ITGA1 Supplemental Figure 1
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
jpeg("ITGA1_interface.jpeg", units="in", width=6, height=5, res=300)

FeaturePlot  (object = Trophoblasts, features = c("ITGA1"),     min.cutoff = 0, order = TRUE, cols = c("grey", "red") , pt.size = 1                ) + NoLegend()
dev.off()



```

#Notch1 Supplemental Figure 1
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
jpeg("NOTCH1_interface.jpeg", units="in", width=6, height=5, res=300)

FeaturePlot  (object = Trophoblasts, features = c("NOTCH1"),     min.cutoff = 0, order = TRUE, cols = c("grey", "red") , pt.size = 1                ) + NoLegend()
dev.off()



```

#ITGA5 Supplemental Figure 1
```{r}
DefaultAssay (object = Trophoblasts) <- "RNA"
jpeg("ITGA5_interface.jpeg", units="in", width=6, height=5, res=300)

FeaturePlot  (object = Trophoblasts, features = c("ITGA5"),     min.cutoff = 0, order = TRUE, cols = c("grey", "red") , pt.size = 1                ) + NoLegend()
dev.off()



```

#Supplemental Figure2 


#Monocle3!

```{r}
detach("package:monocle", unload = TRUE)


library(monocle3)
```

```{r}



cds <- as.cell_data_set      (Trophoblasts, assay = "RNA")
cds <- estimate_size_factors (cds)
cds <- cluster_cells         (cds,
                              reduction_method = c("UMAP"),
                              cluster_method   = "leiden",
                              partition_qval   = 0.05,
                              resolution       = 0.00025)

 

p1  <- plot_cells            (cds,                               show_trajectory_graph = FALSE)
p2  <- plot_cells            (cds, genes=c("cpna-2", "egl-21", "ram-2", "inos-1"))


 plot_cells            (cds, genes=c("CDH1"))


plot_cells   (cds, color_cells_by = "orig.ident", label_cell_groups=FALSE,show_trajectory_graph = FALSE)

marker_test_res <- top_markers(cds, group_cells_by="seurat_clusters", 
                               reference_cells=1000, cores=4)

top_specific_markers <- marker_test_res %>%
                            filter(fraction_expressing >= 0.10) %>%
                            group_by(cell_group) %>%
                            top_n(3, pseudo_R2)

top_specific_marker_ids <- unique(top_specific_markers %>% pull(gene_id))

View(top_specific_marker_ids)


rowData(cds)$gene_short_name <- cds@rowRanges@partitioning@NAMES

rowData(cds)$gene_short_name <- row.names(rowData(cds))



View(rowData(cds)$gene_short_name)

plot_genes_by_group(cds,
                    top_specific_marker_ids,
                    group_cells_by="cluster_name",
                    ordering_type="cluster_row_col",
                    max.size=3)

View(cds)

wrap_plots (p1, p2)
```

```{r}
cds <- learn_graph (cds,
                    use_partition = FALSE,
                    close_loop    = TRUE)
```

```{r}
cds <- order_cells (cds)
```

```{r}
plot_cells (cds,
            label_groups_by_cluster = FALSE,
            label_leaves            = FALSE,
            label_branch_points     = FALSE,
            cell_size                     = 2,
            color_cells_by          = "seurat_clusters")

library(RColorBrewer)

plot_cells (cds,
            color_cells_by                = "pseudotime",
            label_cell_groups             = FALSE,
            label_leaves                  = FALSE,
            label_branch_points           = FALSE,
            graph_label_size              = 1.5,
            cell_size                     = 2,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25)


plot_cells (cds,
            color_cells_by                = "pseudotime",
            label_cell_groups             = FALSE,
            label_leaves                  = FALSE,
            label_branch_points           = FALSE,
            graph_label_size              = 1.5,
            cell_size                     = 2,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25) + NoLegend ()
#Figure 2B

pdf("monocle3.pdf",  width=6, height=5)

plot_cells (cds,
            label_cell_groups             = F,
            label_leaves                  = F,
            label_branch_points           = F,
            color_cells_by                = "ident",
            cell_size                     = 2,
            trajectory_graph_color        = "black",
            trajectory_graph_segment_size = 1.25) 

dev.off()



```

#Supplemental figure 3 - load previously created monocle 2 object
```{r}
detach("package:monocle3", unload = TRUE)


library(monocle)
```

```{r}
#loading my HSMM file
load( "/Volumes/FREECOM HDD/R_ADAM/data/Monocle2_ALLplacentasTrophoblast_3May2021.RData")
```

```{r}

HSMM_genes1 <- row.names(subset(fData(HSMM),
          gene_short_name %in% c("HLA-G","ERVFRD-1","TEAD4")))


HSMM_genes2 <- row.names(subset(fData(HSMM),
          gene_short_name %in% c("MMP14","ADAMTS6", "ADAMTS1")))

HSMM_genes3 <- row.names(subset(fData(HSMM),
          gene_short_name %in% c("MMP15", "ADAM17","ADAM10")))

HSMM_genes4 <- row.names(subset(fData(HSMM),
          gene_short_name %in% c("MMP23B","TLL1", "ADAMTS19")))

HSMM_genes5 <- row.names(subset(fData(HSMM),
          gene_short_name %in% c("TLL2", "MMP19","ADAM10")))


jpeg("Monocle2_CTB_branched_pseudotime.jpeg",  width=3, height=3,units="in", res=300)
plot_genes_branched_pseudotime(HSMM[HSMM_genes1,],
                       branch_point = 1,
                       color_by = "cluster_name",
                       ncol = 1)+ scale_color_manual(breaks = c("CTB 1", "CTB 2", "CTB 3",'CTB 4','cCTB 1' ,'EVT', "SCTp","cCTB 2"), values=c('#CD9600',"#7CAE00","#00BE67",'#00BFC4','#00A9FF','#FF61CC','#F8766D', '#C77CFF'))

dev.off()
jpeg("Monocle2_CTB_branched_pseudotime2.jpeg",  width=3, height=3,units="in", res=300)
plot_genes_branched_pseudotime(HSMM[HSMM_genes2,],
                       branch_point = 1,
                       color_by = "cluster_name",
                       ncol = 1)+ scale_color_manual(breaks = c("CTB 1", "CTB 2", "CTB 3",'CTB 4','cCTB 1' ,'EVT', "SCTp","cCTB 2"), values=c('#CD9600',"#7CAE00","#00BE67",'#00BFC4','#00A9FF','#FF61CC','#F8766D', '#C77CFF'))

dev.off()
jpeg("Monocle2_CTB_branched_pseudotime3.jpeg",  width=3, height=3,units="in", res=300)
plot_genes_branched_pseudotime(HSMM[HSMM_genes3,],
                       branch_point = 1,
                       color_by = "cluster_name",
                       ncol = 1)+ scale_color_manual(breaks = c("CTB 1", "CTB 2", "CTB 3",'CTB 4','cCTB 1' ,'EVT', "SCTp","cCTB 2"), values=c('#CD9600',"#7CAE00","#00BE67",'#00BFC4','#00A9FF','#FF61CC','#F8766D', '#C77CFF'))

dev.off()
jpeg("Monocle2_CTB_branched_pseudotime4.jpeg",  width=3, height=3,units="in", res=300)
plot_genes_branched_pseudotime(HSMM[HSMM_genes4,],
                       branch_point = 1,
                       color_by = "cluster_name",
                       ncol = 1)+ scale_color_manual(breaks = c("CTB 1", "CTB 2", "CTB 3",'CTB 4','cCTB 1' ,'EVT', "SCTp","cCTB 2"), values=c('#CD9600',"#7CAE00","#00BE67",'#00BFC4','#00A9FF','#FF61CC','#F8766D', '#C77CFF'))

dev.off()

jpeg("Monocle2_CTB_branched_pseudotime5.jpeg",  width=3, height=3,units="in", res=300)
plot_genes_branched_pseudotime(HSMM[HSMM_genes5,],
                       branch_point = 1,
                       color_by = "cluster_name",
                       ncol = 1)+ scale_color_manual(breaks = c("CTB 1", "CTB 2", "CTB 3",'CTB 4','cCTB 1' ,'EVT', "SCTp","cCTB 2"), values=c('#CD9600',"#7CAE00","#00BE67",'#00BFC4','#00A9FF','#FF61CC','#F8766D', '#C77CFF'))

dev.off()

```

#Supplemental figure 4 is included in figure 6 analysis





